### 2024-10-25 NEURIPS24/neurips_2024_to_submit.tex

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2024

% ready for submission
\usepackage[final]{neurips_2024}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2024}

% to compile a camera-ready version, add the [final] option, e.g.:
%     \usepackage[final]{neurips_2024}

% to avoid loading the natbib package, add option nonatbib:
%    \usepackage[nonatbib]{neurips_2024}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{wrapfig}
\usepackage[capitalize,noabbrev]{cleveref}

\title{Shapes analysis for time series}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Thibaut Germain$^1$\thanks{Corresponding author. Contact at \texttt{thibaut.germain@ens-paris-saclay.fr}}\\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  \And
  Samuel Gruffaz$^1$\\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  \AND
  Charles Truong$^1$\\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  \And
  Laurent Oudre$^1$\\
  Centre Borelli, ENS Paris-Saclay\\
  4 av. des sciences, 91190 \\
  \And
  Alain Durmus \\
  CMAP, CNRS, Ecole polytechnique\\
  Institut Polytechnique de Paris\\
  91120 Palaiseau, France \\
}

\input{header.tex}
\input{def.tex}

\newcommand\sam[1]{\textcolor{blue}{#1}}
\newcommand\thi[1]{\textcolor{orange}{#1}}

\begin{document}
\footnotetext[1]{Complete affiliation: Université Paris Saclay, Université Paris Cité, ENS Paris Saclay, CNRS, SSA, INSERM, Centre
Borelli, F-91190, Gif-sur-Yvette, France.}
\maketitle

\begin{abstract}
Analyzing inter-individual variability of physiological functions is particularly appealing in medical and biological contexts to describe or quantify health conditions. Such analysis can be done by comparing individuals to a reference one with time series as biomedical data.
This paper introduces an unsupervised representation learning (URL) algorithm for time series tailored to inter-individual studies. The idea is to represent time series as deformations of a reference time series. The deformations are diffeomorphisms parameterized and learned by our method called TS-LDDMM. Once the deformations and the reference time series are learned, the vector representations of individual time series are given by the parametrization of their corresponding deformation. At the crossroads between URL for time series and shape analysis, the proposed algorithm handles irregularly sampled multivariate time series of variable lengths and provides shape-based representations of temporal data.
In this work, we establish a representation theorem for the graph of a time series and derive its consequences on the LDDMM framework. We showcase the advantages of our representation compared to existing methods using synthetic data and real-world examples motivated by biomedical applications.
\end{abstract}

\input{paper.tex}
\begin{ack}
This work was supported by grants from Région Ile-de-France (DIM MathInnov).
Charles Truong is funded by the PhLAMES chair of ENS Paris-Saclay.
\end{ack}

\bibliographystyle{plain}
\bibliography{ref.bib}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\appendix

\input{appendix.tex}
\input{checklist.tex}

\end{document}